---
title: "Human Activity Recognition Using Smartphones Dataset"
author: "Dickson"
date: "2/2/2021"
output:
  md_document:
      variant: markdown_github
---

This study summarizes the means and standard deviations (std) of each variables
according to each activity and each subject prior to identify the most significant variable for predicting human activity.

## Procedure:

1. Training data and testing data for both signals(x) and activities(y), along with the subject labels were merged.
```{r dim}
xtrain <- read.table("./data/UCI HAR Dataset/train/X_train.txt")
xtest <- read.table("./data/UCI HAR Dataset/test/X_test.txt")
ytrain <- read.table("./data/UCI HAR Dataset/train/y_train.txt")
ytest <- read.table("./data/UCI HAR Dataset/test/y_test.txt")
trainsubj <- read.table(file("./data/UCI HAR Dataset/train/subject_train.txt"))
testsubj <- read.table(file("./data/UCI HAR Dataset/test/subject_test.txt"))
train <- cbind(trainsubj, xtrain, ytrain)
test <- cbind(testsubj, xtest, ytest)
data <- rbind(train, test)
dim(data)
```

2. features.txt was parsed to get the variable names of mean and std.
```{r feature}
file <- file("./data/UCI HAR Dataset/features.txt")
features <- readLines(file)
close(file)
length(features)
```

3. mean and std variables extracted from the data set.
```{r meanstd}
target <- features[grepl("mean[(]|std[(]", features)]
length(target)
```

4. Signal variables were renamed and improved based on features.txt
```{r colnames}
library(readr)
index <- parse_number(target)+1
msdata <- data[, c(1, index, ncol(data))]
variables <- gsub("[-0-9() ]", "", target)
variables <- gsub("mean", "MEAN", variables)
variables <- gsub("std", "STD", variables)
names(msdata) <- c("Subject_ID", variables, "Activity")
names(msdata)
```

5. Renaming the descriptive activity names (6 levels of activities)
```{r levels}
labelfile <- readLines("./data/UCI HAR Dataset/activity_labels.txt")
labels <- sub("^[0-9] ", "", labelfile)
msdata$Activity <- factor(msdata$Activity, levels = c(1:6), labels = labels)
levels(msdata$Activity)
```

6. Summarizing dataset with the average of each variable for each activity and each subject
```{r result}
library(dplyr)
df <- tbl_df(msdata)
result <- df %>%
          group_by(Activity, Subject_ID) %>%
          summarize(across(tBodyAccMEANX:fBodyBodyGyroJerkMagSTD, mean)) %>%
          print
```

==================================================================

This project includes the following files:  
- 'README.md'  
- 'CodeBook.md' - Shows information about the variables used.  
- 'run_analysis.R' - Script that can reproduce the data cleaning process.
- 'meanbyActivitySubject.txt' - Output table

==================================================================


The detailed information of the original project are as followed:  

==================================================================  

Human Activity Recognition Using Smartphones Dataset
Version 1.0

==================================================================  
Jorge L. Reyes-Ortiz, Davide Anguita, Alessandro Ghio, Luca Oneto.
Smartlab - Non Linear Complex Systems Laboratory
DITEN - Università degli Studi di Genova.
Via Opera Pia 11A, I-16145, Genoa, Italy.
activityrecognition@smartlab.ws
www.smartlab.ws
==================================================================  

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

For each record it is provided:

- Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
- Triaxial Angular velocity from the gyroscope. 
- A 561-feature vector with time and frequency domain variables. 
- Its activity label. 
- An identifier of the subject who carried out the experiment.


License:
Use of this dataset in publications must be acknowledged by referencing the following publication [1] 

[1] Davide Anguita, Alessandro Ghio, Luca Oneto, Xavier Parra and Jorge L. Reyes-Ortiz. Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine. International Workshop of Ambient Assisted Living (IWAAL 2012). Vitoria-Gasteiz, Spain. Dec 2012

This dataset is distributed AS-IS and no responsibility implied or explicit can be addressed to the authors or their institutions for its use or misuse. Any commercial use is prohibited.

Jorge L. Reyes-Ortiz, Alessandro Ghio, Luca Oneto, Davide Anguita. November 2012.
